### 2019-08-28 manuscript/manuscript.tex
ms changes

% TODO
% - generate figures and write about them

@@

We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
ten replicate populations exposed to a high temperature environment and evolved
for 60 generations, and sequenced every ten generations. Using the seven
timepoints and ten replicate populations, we estimated a bias-corrected $60
\times 60$ temporal-replicate variance-covariance matrix (see XXX for details).
Since each replicate population was sequenced every ten generations, the
timepoints $t_0 = 0$ generations, $t_1 = 10$ generations, $t_2 = 20$
generations, etc., lead to observed allele frequency changes across ten
generation chunks, $\Delta p_{t_0}, \Delta p_{t_1}, \ldots, \Delta p_{t_6}$.
Consequently, the ten temporal covariance matrices for each of the ten
replicate populations have off-diagonal elements of the form $\cov(\Delta
p_{t_0}, \Delta p_{t_1}) = \cov(p_{t_1} - p_{t_0}, p_{t_2} - p_{t_1}) =
\sum_{i=0}^{10} \sum_{j=10}^{20} \cov(\Delta p_i, \Delta p_j)$. Each diagonal
element has the form $\var(\Delta p_{t_0}) = \sum_{i=0}^{t_0} \var(\Delta
p_{i}) + \sum_{i \ne j}^{t_0} \cov(\Delta p_{i}, \Delta p_{j})$, and is thus a
combination of the effects of drift and selection, as both the variance in
allele frequency changes and cumulative temporal autocovariances terms increase
the variance in allele frequency. With sampling each generation, one could more
accurately partition the total variance in allele frequency change
\parencite{Buffalo2019-io}; while we cannot directly estimate the contribution
of linked selection to the variance in allele frequency change here, the
presence of a positive observed covariance between allele frequency change can
only be caused linked selection.

Averaging across replicate populations, we find positive temporal covariances
that are statistically significant (p < XXX) consistent with linked selection
acting to affect allele frequency changes over very short time periods. We
visualize these covariances in Figure \label{figure-1} (A), which depicts the
temporal covariances through time, for each of the five rows covariance matrix.
Each row represents the temporal covariance $\cov(\Delta p_s, \Delta p_t)$,
between some initial reference generation $s$ (the row of the matrix), and some
later timepoint $t$ (the column of the matrix). For each row, the covariances
at first are positive, and then decay towards zero as expected when directional
selection affects linked variants' frequency trajectories until ultimately
linkage disequilibrium and additive genetic variance for decay
\parencite{Buffalo2019-io}. Note that per replicate, the signal is a bit
noisier; see Supplementary Figures XXX.

While the presence of positive temporal covariances is consistent with linked
selection affecting allele frequencies over time, this measure not easily
interpretable. Alternatively, we can quantify the impact of linked selection
through the ratio of total covariance in allele frequency change to the total
variance in allele frequency change. Since the total variation in allele
frequency change can be decomposed into variance and covariance components,
$\var(p_t - p_0) = \sum_{i \ne j} \cov(\Delta p_i, \Delta p_j) / \var(p_t -
p_0)$, and the covariances are zero when only drift acts, this is a
conservative measure of how much of the variance in allele frequency change is
caused by linked selection \parencite{Buffalo2019-io}. Since timepoints every
ten generations were sequenced, we cannot directly estimate the temporal
covariances

>>>>

% TODO
% - generate figures and write about them
% - Figure 2 has some overplotting issues.

@@

We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
ten replicate populations exposed to a high temperature environment and evolved
for 60 generations, and sequenced every ten generations. Using the seven
timepoints and ten replicate populations, we estimated a bias-corrected $60
\times 60$ temporal-replicate variance-covariance matrix (see XXX for details).
Since each replicate population was sequenced every ten generations, the
timepoints $t_0 = 0$ generations, $t_1 = 10$ generations, $t_2 = 20$
generations, etc., lead to observed allele frequency changes across ten
generation blocks, $\Delta p_{t_0}, \Delta p_{t_1}, \ldots, \Delta p_{t_6}$.
Consequently, the ten temporal covariance matrices for each of the ten
replicate populations have off-diagonal elements of the form $\cov(\Delta
p_{t_0}, \Delta p_{t_1}) = \cov(p_{t_1} - p_{t_0}, p_{t_2} - p_{t_1}) =
\sum_{i=0}^{10} \sum_{j=10}^{20} \cov(\Delta p_i, \Delta p_j)$. Each diagonal
element has the form $\var(\Delta p_{t_0}) = \sum_{i=0}^{t_0} \var(\Delta
p_{i}) + \sum_{i \ne j}^{t_0} \cov(\Delta p_{i}, \Delta p_{j})$, and is thus a
combination of the effects of drift and selection, as both the variance in
allele frequency changes and cumulative temporal autocovariances terms increase
the variance in allele frequency. With sampling each generation, one could more
accurately partition the total variance in allele frequency change
\parencite{Buffalo2019-io}; while we cannot directly estimate the contribution
of linked selection to the variance in allele frequency change here, the
presence of a positive observed covariance between allele frequency change can
only be caused linked selection. Overall, we expect observed temporal
covariances between timepoints to be weak, some fraction of the additive
genetic variance and linkage disequilibrium that creates temporal
autocovariance decays in the ten generations between sequencing.

Averaging across replicate populations, we find positive temporal covariances
across time that are statistically significant (p < XXX) consistent with linked
selection acting to affect allele frequency changes over very short time
periods. We visualize these covariances in Figure \label{figure-1} (A), which
depicts the temporal covariances through time, for each of the five rows
covariance matrix.  Each row represents the temporal covariance $\cov(\Delta
p_s, \Delta p_t)$, between some initial reference generation $s$ (the row of
the matrix), and some later timepoint $t$ (the column of the matrix). For each
row, the covariances at first are positive, and then decay towards zero as
expected when directional selection affects linked variants' frequency
trajectories until ultimately linkage disequilibrium and additive genetic
variance for decay \parencite{Buffalo2019-io}. Note that per replicate, the
signal is a bit noisier; see Supplementary Figures XXX.

While the presence of positive temporal covariances is consistent with linked
selection affecting allele frequencies over time, this measure not easily
interpretable. Additionally, we can quantify the impact of linked selection on
allele frequency change as the ratio of total covariance in allele frequency
change to the total variance in allele frequency change. Since the total
variation in allele frequency change can be decomposed into variance and
covariance components, $\var(p_t - p_0) = \sum_{i \ne j} \cov(\Delta p_i,
\Delta p_j) / \var(p_t - p_0)$, and the covariances are zero when drift acts
alone, this is a lower bound on how much of the variance in allele frequency
change is caused by linked selection \parencite{Buffalo2019-io}. We call this
measure $G(t)$, which is the total effect of linked selection between the
initial generation $0$ and some later generation $t$, which can be varied to
see how this quantity grows through time. As with the temporal covariances, the
study design of \textcite{Barghi2019} leads our measure $G(t)$ to be even more
conservative, since the temporal covariances with each ten-generation block
between sequenced timepoints are not directly observable, and are not included
in the numerator of $G(t)$. Still, we find a remarkably strong signal

% NOTE in supp. material why the average replicate G(t) looks lower than
% replicates.
